### 2017-06-15 manuscript/manuscript.Rmd
Edit # Version Control Systems

Version Control Systems (VCS; also known as Source Control systems) are computer programs designed for tracking changes to computer code, and sharing the code---and its history---with others. Although initially developed for writing computer code collaboratively, it is easy to recognize the benefits of VCS for scientific production and collaboration as well. For example, behavioral experiments are often written in a programming language, and can include multiple authors and versions. Keeping track of the versions of the program, and allowing many authors to contribute to it (without breaking the code) are problems that VCSs were specifically designed for. By extension, VCS easily adopts to other aspects of the research cycle, such as curating data across computers, laboratories, and experiments; even writing manuscripts is easier when it is recognized that manuscripts (usually) undergo multiple rounds of revision, and that multiple authors might want to keep track of all these different versions.

@@

<!-- Examples of problems in psychology solved by VCS somewhere here. -->

\noindent
\fbox{\begin{minipage}{\textwidth}
  \begin{itemize}
    \item Problem: Researchers usually want to try different ways of visualizing and modeling their data, and sometimes it is difficult to keep track of all the different versions.
      \begin{itemize}
        \item Solution: VCS allows researchers to save each version of the analysis into a history from where it can be easily recalled. This approach also eschews the need for creating multiple copies of the same analysis files with small tweaks, as illustrated \href{http://phdcomics.com/comics/archive.php?comicid=1531}{here}.
      \end{itemize}
  \end{itemize}
  \begin{itemize}
    \item Problem:
      \begin{itemize}
        \item Solution:
      \end{itemize}
  \end{itemize}
\end{minipage}}

@@

It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git[^git-name] is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook. It can be freely downloaded at <https://git-scm.com/>. Before we explain the basics of using Git, we present brief installation instructions for Windows and Apple users.

@@

@@

[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.

>>>>

Version Control Systems (VCS; also known as Source Control Systems) are computer programs designed for tracking changes to computer code, and sharing the code with others. Although initially developed for writing computer code collaboratively[^vcs-collaborate], it is easy to recognize the benefits of VCS for scientific production and collaboration as well. For example, behavioral experiments are often written in a programming language, and can have multiple authors and versions. Keeping track of the versions of the program, and allowing many authors to contribute to it (without breaking the code) are problems that VCSs were specifically designed for. By extension, VCS is easily adopted to other aspects of the research cycle, such as curating data across computers and laboratories, or writing manuscripts with multiple authors, versions, and sources of results.

<!-- Examples of problems in psychology solved by VCS. -->
\noindent
\fbox{\begin{minipage}{\textwidth}
  \begin{itemize}
    \item Problem: Researchers want to try different ways of visualizing and modeling their data, and keep track of all the different versions they try.
      \begin{itemize}
        \item Solution: VCS allows researchers to save each version of the analysis into a history from where it can be easily recalled. This approach also eschews the need for creating multiple copies of the same analysis files with small tweaks, as illustrated \href{http://phdcomics.com/comics/archive.php?comicid=1531}{here}.
      \end{itemize}
  \end{itemize}
  \begin{itemize}
    \item Problem:
      \begin{itemize}
        \item Solution:
      \end{itemize}
  \end{itemize}
\end{minipage}}

@@

@@

It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git[^git-name] is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others). It can be freely downloaded at <https://git-scm.com/>. Before we explain the basics of using Git, we present brief installation instructions for Windows and Apple users.

@@

- Born open data [@rouder_what_2015]

@@

[^git-version]: As of the writing of this article, the current version of Git is 2.13.1.
[^git-name]: The creator of Git, Linus Torvalds, named Git after himself as "the stupid content tracker" (@mcmillan_after_2005; <https://git-scm.com/docs/git.html>)
[^simple]: By "simple", we only mean that the technical aspects of the project, related to curating materials related to it, are simple. We of course do not suggest that there might be anything theoretically or scientifically simple in such projects.
[^vcs-collaborate]: The software we present below is used used by major software developers such as Microsoft, Google, and Facebook on code bases with hundreds of contributors.
